MplusAutomation
===============

The `MplusAutomation` package leverages the flexibility of the `R`
language to automate latent variable model estimation and
interpretation using Mplus, a powerful latent variable modeling
program developed by Muthén and Muthén
(www.statmodel.com). Specifically, `MplusAutomation` provides routines
for creating related groups of models, running batches of models, and
extracting and tabulating model parameters and fit statistics.

Installation
------------

You can install the latest release of `MplusAutomation` directly from
[CRAN](http://cran.r-project.org/package=MplusAutomation) by running

```r
install.packages("MplusAutomation")
```

Alternately, if you want to try out the latest development
`MplusAutomation` code, you can install it straight from github using
Hadley Wickham's `devtools` package. If you do not have `devtools`
installed, first install it and then install `MplusAutomation`.

```r
#install.packages("devtools")
require(devtools)

install_github("MplusAutomation", "michaelhallquist")
```

Questions
---------

For questions, answers, and updates on the status of the
`MplusAutomation` package, email or subscribe to the Google group
[list](https://groups.google.com/forum/#!forum/mplusautomation).

Examples
--------

You can find a detailed example of how to use the `MplusAutomation`
package in the
[vignette](https://github.com/michaelhallquist/MplusAutomation/blob/master/inst/doc/Vignette.pdf)
.


Here is an example of using the package to run a simple path model
using the `mtcars` dataset built into `R`.

```{r}
require(MplusAutomation)

pathmodel <- mplusObject(
   TITLE = "MplusAutomation Example - Path Model;",
   MODEL = "
     mpg ON hp;
     wt ON disp;",
   rdata = mtcars)

fit <- mplusModeler(pathmodel, "mtcars.dat", modelout = "model1.inp", run = 1L)
```

That is all it takes to run Mplus!  `MplusAutomation` takes care of
figuring out which variables from your `R` dataset are used in the
model and which are not (if it get's confused, you can also specify
`usevariables`).  It creates a dataset suitable for Mplus, calls Mplus
to run the model on the dataset, and reads it back into `R`.

There is even pretty printing now.  To see the results:

```{r, include=FALSE}
require(texreg)
```

<pre><code>
```{r, results = 'asis'}
require(texreg)
screenreg(extract(fit, summaries = c("Observations", "CFI", "SRMR")), single.row=TRUE)
```
</code></pre>

The fit is not great, to add some extra paths we can update the model.

```{r}
pathmodel2 <- update(pathmodel, MODEL = ~ . + "
    mpg ON disp;
    wt ON hp;")

fit2 <- mplusModeler(pathmodel2, "mtcars2.dat", modelout = "model2.inp", run = 1L)
```

We can make some pretty output of both models:

<pre><code>
```{r, results = 'asis'}
screenreg(list(
  extract(fit, summaries = c("Observations", "CFI", "SRMR")),
  extract(fit2, summaries = c("Observations", "CFI", "SRMR"))),
  single.row=TRUE)
```
</code></pre>

How to Help
-----------

If you have a tutorial or examples using `MplusAutomation`, please add them to
the github [Wiki](https://github.com/michaelhallquist/MplusAutomation/wiki).

In addition, on the [Wiki](https://github.com/michaelhallquist/MplusAutomation/wiki),
is a list of publications that cite or use `MplusAutomation`.  If you use `MplusAutomation`
in your own work --- papers, posters, presentations, etc. --- please add a citation to the
list, and if possible, include an abstract or link to the full text. This helps us get to
know our users and how `MplusAutomation` is being used.

Finally, if you find bugs or have suggestions for new features or ways to enhance
`MplusAutomation`, please let us know!  Just click the 'Issues' button at the top
of the github page or go
[here](https://github.com/michaelhallquist/MplusAutomation/issues?state=open)
and open a New Issue.

Lastly, if you use `MplusAutomation` and have space, we greatly appreciating citations.
In addition to being easier to track, the recognition and credit help make it
easier for us to continue putting our time into developing and sharing this software!




